---
title: "Data transformation with dplyr"
subtitle: "Practice with penguins"
format: html
editor_options: 
  chunk_output_type: console
editor: visual
---

```{r}
#| label: load-packages
library(palmerpenguins)
library(dplyr)
```

All exercises in this assignment use the `penguins` data as a starting point. 

1. Run the code chunk that contains `glimpse(penguins)`.
2. How many variables are in the dataset?
3. How many observations are in the dataset?

```{r}
glimpse(penguins)
```


## Task 1: Create a subset of the data using filter()

Use `filter()` to create a subset from `penguins` that only contains observations for Adelie penguins.

```{r}

```

Use `filter()` to create a subset from `penguins` that only contains observations where body mass is less than or equal to 2900 g.

```{r}

```

Use `filter()` to create a subset from `penguins` that only contains observations for Adelie penguins with a bill length greater than 40 mm.

```{r}

```

Use `filter()` to create a subset from `penguins` that excludes observations for chinstraps.

```{r}

```

Use `filter()` to create a subset from `penguins` that only contains gentoo penguins with a bill depth greater than or equal to 15.5 millimeters.
```{r}

```


Use `filter()` to create a subset from `penguins` that contains observations for male penguins recorded at Dream and Biscoe Islands.

Use `filter()` to create a subset from `penguins` that contains observations for female Adelie penguins with bill lengths less than 35 mm.

Use `filter()` to create a subset from `penguins` containing observations for female chinstrap penguins on Dream and Torgersen Islands.

Use `filter()` to create a subset from `penguins` that contains penguins that are either gentoos OR have a body mass greater than 4500 g.


## Task 2: Add new columns with mutate()

Add a column to `penguins` that contains a new column `flipper_m`, which is the `flipper_length_mm` (flipper length in millimeters) converted to units of meters.

```{r}

```

Add a new column to `penguins` that contains a new column `body_mass_kg`, which is the `body_mass_g` (body mass in grams) converted to units of kilograms.

```{r}

```

Add a new column to `penguins` that contains a new column `bill_ratio`, which is the ratio of bill length to bill depth.

```{r}

```

Add a new column to `penguins` with a sequence of values from 1 to the length of the data frame.

```{r}

```

## Task 3: Summarize data with group_by() and summarize() & count()

Starting with `penguins`, group the data by species , then create a summary table containing the maximum and minimum length of flippers (call the columns flip_max and flip_min). How will you handle NA values?

```{r}
penguins |> 
  group_by(species) |> 
  summarize(flip_max = max(flipper_length_mm), 
            flip_min = min(flipper_length_mm))

```

Starting with `penguins`, group the data by species and year, then create a summary table containing the mean bill depth (call this bill_depth_mean) and mean bill length (call this bill_length_mean) for each group.

```{r}

penguins |> 
  group_by(species, year) |> 
  summarize(bill_depth_mean = mean(bill_depth_mm), 
            bill_length_mean = mean(bill_length_mm))


```

Use the `count()` function to count the number of observations for each species in `penguins`.

```{r}

```

Use the `count()` function to count the number of observations for each species and island in `penguins`.

```{r}


```

